The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2017-01-21 script/en/chapters/movies.tex
CT Book ed 12: Deleted some comments; added Python to Index.

	Poe not only was a well-known writer, but also a talented cryptographer.
	His story is also told in the book {\em Code Breaking} \cite{Kippenhahn1997}.
    }\\
    In this short story Poe tells as first-person narrator about his
    acquaintanceship with the curious Mr. Legrand. They detect the fabulous
    treasure of captain Kidd via a gold bug and a vellum found at the coast
    of New England.\\
    The cipher consists of 203 cryptic symbols and it proves to be a
    general monoalphabetic substitution cipher (see
    chapter~\ref{monoalphabeticSubstitutionCiphers}).
    The story tells how they solve the riddle step by step using a combination
    of semantic and syntax analysis (frequency analysis of single letters in
    English texts).\footnote{
      Part of the learning material above is a Python program which can be used
      to decrypt the ciphertext with Python {\em and} with SageMath\index{SageMath}.
      See the code example in \ref{s:appendix-Code-for-light-fiction-books}.
    }\\
    In this novel the code breaker Legrand says the famous statement:
    ``Yet it may be roundly asserted that human ingenuity cannot concoct a
    cipher which human ingenuity cannot resolve -- given the according
    dedication.''\\
    % Yet it may be roundly asserted that human ingenuity cannot concoct a
    % cipher which human ingenuity cannot resolve...

@@

Using the following Python code\footnote{%
  See \cite{Witten1998} and the file \verb|2_MonoKidd.zip| at
  \url{http://bscw.schule.de/pub/bscw.cgi/159132}.
}
you can decrypt the ciphertext of Captain Kidd (see the original text of
``The Gold Bug'' in \url{http://pinkmonkey.com/dl/library1/gold.pdf},
page 21) with Python or with SageMath\index{SageMath}.

@@

\end{Verbatim}
\caption{Decryption of the Gold-Bug ciphertext from the novel of E.A. Poe (with Python)}
\label{Lit_Python-sample_Gold-bug}
\end{sagecode}

@@

\paragraph*{Remark 2:}
The code uses the Python functions ``maketrans'' und ``translate'' from the
String package (see e.g. the Python 2.7 documentation, section ``functions'').

@@

\paragraph*{Remark 3:}
The following screenshot shows the execution of the Python code on the SageMathCell
server (\url{http://sagecell.sagemath.org/}).

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.53]{figures/Using-SageMathCell-with-Python-for-Poes-GoldBug}
\caption{Usage of SageMathCell to decipher Poe's Gold Bug (with Python)} 
\label{MOV_Using-SageMathCell-with-Python-for-Poes-GoldBug}
\end{center}
\end{figure}

@@

\noindent It's evident how less code is needed with Python or Sage for such tasks.
In the above sample there were 2 unnecessary comment lines, 3 lines for input,
3 lines of real code, and 3 lines for the output; really necessary were only
7 lines of code.

>>>>

	Poe not only was a well-known writer, but also a talented cryptographer.
	His story is also told in the book {\em Code Breaking} \cite{Kippenhahn1997}.
    }\\
    In this short story Poe tells as first-person narrator about his
    acquaintanceship with the curious Mr. Legrand. They detect the fabulous
    treasure of captain Kidd via a gold bug and a vellum found at the coast
    of New England.\\
    The cipher consists of 203 cryptic symbols and it proves to be a
    general monoalphabetic substitution cipher (see
    chapter~\ref{monoalphabeticSubstitutionCiphers}).
    The story tells how they solve the riddle step by step using a combination
    of semantic and syntax analysis (frequency analysis of single letters in
    English texts).\footnote{
      Part of the learning material above is a Python\index{Python} program which can be used
      to decrypt the ciphertext with Python {\em and} with SageMath\index{SageMath}.
      See the code example in \ref{s:appendix-Code-for-light-fiction-books}.
    }\\
    In this novel the code breaker Legrand says the famous statement:
    ``Yet it may be roundly asserted that human ingenuity cannot concoct a
    cipher which human ingenuity cannot resolve -- given the according
    dedication.''\\
    % Yet it may be roundly asserted that human ingenuity cannot concoct a
    % cipher which human ingenuity cannot resolve...

@@

Using the following Python\index{Python} code\footnote{%
  See \cite{Witten1998} and the file \verb|2_MonoKidd.zip| at
  \url{http://bscw.schule.de/pub/bscw.cgi/159132}.
}
you can decrypt the ciphertext of Captain Kidd (see the original text of
``The Gold Bug'' in \url{http://pinkmonkey.com/dl/library1/gold.pdf},
page 21) with Python or with SageMath\index{SageMath}.

@@

\end{Verbatim}
\caption{Decryption of the Gold-Bug ciphertext from the novel of E.A. Poe (with Python\index{Python})}
\label{Lit_Python-sample_Gold-bug}
\end{sagecode}

@@

\paragraph*{Remark 2:}
The code uses the Python\index{Python} functions ``maketrans'' und ``translate'' from the
String package (see e.g. the Python 2.7 documentation, section ``functions'').

@@

\paragraph*{Remark 3:}
The following screenshot shows the execution of the Python\index{Python} code on the SageMathCell
server (\url{http://sagecell.sagemath.org/}).

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.53]{figures/Using-SageMathCell-with-Python-for-Poes-GoldBug}
\caption{Usage of SageMathCell to decipher Poe's Gold Bug (with Python\index{Python})} 
\label{MOV_Using-SageMathCell-with-Python-for-Poes-GoldBug}
\end{center}
\end{figure}

@@

\noindent It's evident how less code is needed with Python\index{Python} or Sage for such tasks.
In the above sample there were 2 unnecessary comment lines, 3 lines for input,
3 lines of real code, and 3 lines for the output; really necessary were only
7 lines of code.

### 2017-01-21 script/en/chapters/numbertheory.tex
CT Book ed 12: Deleted some comments; added Python to Index.

When Lenstra et al published their paper \cite{Lenstra2012} in Feb 2012, they did not publish the source code. However, soon afterwards the source code of a similar program was published at the CrypTool website\footnote{%
\url{http://www.cryptool.org/en/ctp-dokumentation-en/361-ctp-paper-rsa-moduli}
}
in Python and C++, and -- again a bit later -- at the page used by \cite{Heninger2012}\footnote{\url{https://www.factorable.net/}
}.
The fastest code known to me comes with \cite{Heninger2012}.

@@

The results of modular arithmetic are used extensively in\index{cryptography!modern}
modern cryptography. Here we will provide a few examples from cryptography
using small\footnote{%
  In the RSA procedure, we call numbers ``small'' if the bit lengths are
  much less than $1024$ bits (i.e. $308$ decimal points). In practice,
  $1024$ bits is currently considered the minimum length for a secure
  RSA modul.
} numbers.\footnote{%
  Within the series {\em RSA \& Co. at school: Modern cryptology, old
  mathematics, and subtle protocols} there are didactically very well
  prepared articles with code samples in Python and SageMath.
  Unfortunately these are currently only available in German.
  See for instance \cite{Witten2015}.\index{RSA \& Co. at school}
}

@@

   \item SageMath \index{SageMath}\\
         Excellent, open source computer algebra system with Python as script
         language, used to build the code samples in this chapter.
	 See the introduction in chapter~\ref{s:appendix-using-sage}.\\
         \url{http://www.sagemath.org/}\\
         \url{http://en.wikipedia.org/wiki/Sage_%28mathematics_software%29}

>>>>

When Lenstra et al published their paper \cite{Lenstra2012} in Feb 2012, they did not publish the source code. However, soon afterwards the source code of a similar program was published at the CrypTool website\footnote{%
\url{http://www.cryptool.org/en/ctp-dokumentation-en/361-ctp-paper-rsa-moduli}
}
in Python\index{Python} and C++, and -- again a bit later -- at the page used by \cite{Heninger2012}\footnote{\url{https://www.factorable.net/}
}.
The fastest code known to me comes with \cite{Heninger2012}.

@@

The results of modular arithmetic are used extensively in\index{cryptography!modern}
modern cryptography. Here we will provide a few examples from cryptography
using small\footnote{%
  In the RSA procedure, we call numbers ``small'' if the bit lengths are
  much less than $1024$ bits (i.e. $308$ decimal points). In practice,
  $1024$ bits is currently considered the minimum length for a secure
  RSA modul.
} numbers.\footnote{%
  Within the series {\em RSA \& Co. at school: Modern cryptology, old
  mathematics, and subtle protocols} there are didactically very well
  prepared articles with code samples in Python\index{Python} and SageMath.
  Unfortunately these are currently only available in German.
  See for instance \cite{Witten2015}.\index{RSA \& Co. at school}
}

@@

   \item SageMath \index{SageMath}\\
         Excellent, open source computer algebra system with Python\index{Python}
	 as script language, used to build the code samples in this chapter.
	 See the introduction in chapter~\ref{s:appendix-using-sage}.\\
         \url{http://www.sagemath.org/}\\
         \url{http://en.wikipedia.org/wiki/Sage_%28mathematics_software%29}

### 2017-01-21 script/en/chapters/what-is-sage.tex
CT Book ed 12: Deleted some comments; added Python to Index.

This book includes numerous code samples using SageMath. SageMath is an open
source computer algebra system (CAS) that supports teaching, study and
research in mathematics.  It combines many high-quality open
source packages\footnote{%
To get an impression of how big SageMath is: After downloading the source of SageMath 4.1,
it took around 5 hours on an average Linux PC to compile the whole system including
all libraries. The compiled version occupied 1.8 GB disk space.
}
and provides access to their functionalities via a
common interface, namely, a Python\footnote{%
There is also an easy interface to the C language, called Cython, which can be used
to substantially speed up functions in SageMath.\\
See \url{http://openwetware.org/wiki/Open_writing_projects/Sage_and_cython_a_brief_introduction}.
} based programming language.

>>>>

This book includes numerous code samples using SageMath. SageMath is an open
source computer algebra system (CAS) that supports teaching, study and
research in mathematics.  It combines many high-quality open
source packages\footnote{%
To get an impression of how big SageMath is: After downloading the source of SageMath 4.1,
it took around 5 hours on an average Linux PC to compile the whole system including
all libraries. The compiled version occupied 1.8 GB disk space.
}
and provides access to their functionalities via a
common interface, namely, a Python\index{Python}\footnote{%
There is also an easy interface to the C language, called Cython, which can be used
to substantially speed up functions in SageMath.\\
See \url{http://openwetware.org/wiki/Open_writing_projects/Sage_and_cython_a_brief_introduction}.
} based programming language.
